---
title: "P8105 Homework 5"
author: "Maya Krishnamoorthy"
date: "2024-11-13"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

set.seed(1)
```

## Problem 1

Suppose you put 𝑛 people in a room, and want to know the probability that at least two people share a birthday. For simplicity, we’ll assume there are no leap years (i.e. there are only 365 days) and that birthdays are uniformly distributed over the year (which is actually not the case).

```{r}
bday_sim = function(n) {
  
  bdays = sample(1:365, size = n, replace = TRUE)
  
  duplicate = length(unique(bdays)) < n
  
  return(duplicate)
}
```

```{r}
sim_res = 
  expand_grid(
    n = 2:50,
    iter = 1:10000
  ) |> 
  mutate(res = map_lgl(n, bday_sim)) |> 
  group_by(n) |> 
  summarize(prob = mean(res))

sim_res |> 
  ggplot(aes(x = n, y = prob )) + 
  geom_line() +
  labs(
    title = "Probability of Duplicate Birthdays as a Function of Group Size"
  )
```

The probability of having a duplicate birthday increases as the group size increases. For small group sizes, the probability of a duplicate birthday is close to zero. It starts off very low because with fewer people, the chance of having a matching birthday is small. As the group size approaches 50, the probability approaches 1 (or 100%) meaning that it is almost certain that there will be at least one pair of people that share a birthday.


## Problem 2

